Tables and Figures to put in manuscript

### 2023-11-27 Docs/manuscript.Rmd
simulation result summary

---
title: "Dynamic Prediction with Fast Generalized Functional Principal Component analysis"
author: "Ying Jin"
date: "2023-03-07"
output: 
  pdf_document:
    citation_package: natbib
    keep_text: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

---
title: "Tables and Figures to put in manuscript"
author: "Ying Jin"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: yes
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    font: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = "")

set.seed(1120)

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(knitr)
library(mvtnorm)
library(mgcv)
library(splines)
theme_set(theme_minimal())

load(here("Data/SimOutput_fGFPCA.RData"))
load(here("Data/SimOutput_GLMMadaptive.RData"))

## prediction window 
window <- seq(0, 1, by = 0.2)
M <- length(pred_list_all)
```

# Larger-scale simulation

## Figure

```{r fig_pred}
rand_id <- sample(501:600, size = 4)
# range(pred_list_all[[1]]$pred0.2, na.rm = T)
fig_fgfpca <- pred_list_all[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time")

fig_adglmm <- pred_list_ref[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time")

```

```{r}
ggarrange(fig_fgfpca, fig_adglmm,
          nrow = 1, common.legend = T)
```
## ISE

```{r}
## ISE container 
ise_mat <- ise_mat_ref <- array(NA, 
                                dim = c(length(window)-2, length(window)-2, M))
# dims: prediction window, max obs time, simulation iter

```

```{r ise_fGFPCA}
## calculation
for(m in 1:M){
  this_df <- pred_list_all[[m]]
  ise_tb <- pred_list_all[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_mat[, ,m] <- as.matrix(ise_tb)

}

mean_ise <- apply(ise_mat, c(1, 2), mean)

# mean_ise <- data.frame(mean_ise) %>%
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_ise) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r ise_GLMMadaptive}
## calculation
for(m in 1:M){
  this_df <- pred_list_ref[[m]]
  ise_tb <- pred_list_ref[[m]] %>%
    mutate(err1 = (pred0.2-eta_i)^2,
           err2 = (pred0.4-eta_i)^2,
           err3 = (pred0.6-eta_i)^2,
           err4 = (pred0.8-eta_i)^2) %>%
    select(id, t, starts_with("err")) %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    group_by(window, id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) %>% 
    group_by(window) %>% 
    summarize_at(vars(err1, err2, err3, err4), mean) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("err"))
  ise_mat_ref[, ,m] <- as.matrix(ise_tb)

}

mean_ise_ref <- apply(ise_mat_ref, c(1, 2), mean)
# mean_ise_ref <- data.frame(mean_ise_ref) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_ise_ref) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_ise, mean_ise_ref, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

## AUC

```{r auc_func}
## a function to calculate AUC
get_auc <- function(y, pred){
  if(sum(is.na(y))>0 | sum(is.na(pred))>0){
    auc <- NA
  }
  else{
    this_perf <- performance(prediction(pred, y), measure = "auc")
    auc <- this_perf@y.values[[1]]
  }
  return(auc)
}

```

```{r}
## auc container 
auc_mat <- array(NA, dim = c(length(window)-2, length(window)-2, M))
auc_mat_ref <- array(NA, dim = c(length(window)-2, length(window)-2, M))
```

```{r auc_fGFPCA}
for(m in 1:M){
  this_df <- pred_list_all[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_mat[, ,m] <- as.matrix(auc_tb)

}

mean_auc <- apply(auc_mat, c(1, 2), mean)
# mean_auc <- data.frame(mean_auc) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_auc) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r auc_GLMMadaptive}
for(m in 1:M){
  this_df <- pred_list_ref[[m]]
  auc_tb <- this_df %>% 
    mutate(window = cut(t, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred0.2),
              auc2 = get_auc(Y, pred0.4),
              auc3 = get_auc(Y, pred0.6),
              auc4 = get_auc(Y, pred0.8)) %>%
    filter(window != "[0,0.2]") %>% 
    select(starts_with("auc"))
  auc_mat_ref[, ,m] <- as.matrix(auc_tb)

}

mean_auc_ref <- apply(auc_mat_ref, c(1, 2), mean)
# mean_auc <- data.frame(mean_auc) %>% 
#   mutate(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
#          .before = 1)
colnames(mean_auc_ref) <- c("0.2", "0.4", "0.6", "0.8")
```

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_auc, mean_auc_ref, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Area under the ROC curve") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

```{r time}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive"),
  "Fit" = c(mean(fit_time)/60, mean(fit_time_ref)),
           "Prediction"= c(mean(pred_time), mean(pred_time_ref)/60)) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Computation time (minutes)") %>%
  kable_styling(full_width = F) 
```

I think we could say that while the total time spend on model fitting + prediction are similar between two methods, fGFPCA achieved much better flexibility and much better predictive performance of prediction under every scenario.

# Small-scale simulation

# NHANES data application
